Causal Forest Approach for Site-specific Input Management via On-farm Precision Experimentation

### 2022-02-07 Writing/manuscript_cea.rmd
made significant modifications to the manuscript and response to reviewers

```{r echo = F, include = FALSE}
library(knitr)
library(here)
here::i_am("GitControlled/Writing/manuscript_cea.rmd")
# opts_knit$set(root.dir = here())

@@

opts_chunk$set(
  fig.align = "center",
  fig.retina = 5,
  warning = F,
  message = F,
  cache = FALSE, # <-- T
  cache.lazy = FALSE,
  echo = F
)
```

```{r, cache = FALSE}
# === packages ===#
library(tidyverse)
library(data.table)

library(png)
library(RColorBrewer)
library(patchwork)
library(plotly)
library(magrittr)
library(ggplot2)
library(viridis)
library(tmap)
library(ggthemes)
library(mgcv)
library(ggpubr)
library(flextable)
library(officer)
library(officedown)
library(modelsummary)
library(raster)
library(bookdown)
library(png)
library(grid)
```

@@

CF was developed specifically for identifying heterogeneous treatment effects [@athey2016recursive; @Wager2018a]. Like RF, CF works by building an ensemble of trees, called *causal trees* (CT), by recursively partitioning covariate space into two leaves, where each leaf estimates conditional average treatment effects as the mean differences between the dependent variable (here, yield) of the treated and control observations that belong to the leaf (please see Appendix for an example of a causal tree).

@@

Monte Carlo simulations were conducted for nitrogen experiments for corn production to estimate site-specific EONRs. The simulations' first step generated one thousand synthetic fields that "resembled" an actual field. Then, for each simulation round, one of the fields was used as a training data set, and another field that is different from the one used as a training data set in that round was used as a validation data set. Field characteristics and yield were simulated assuming the Mitscherlich-Baule (MB) yield response function. We used the RF, BRF, and CNN ML methods to estimate yield response functions, and we used the CF method to estimate treatment effects. For each method, we estimated site-specific EONRs using the training data set, and checked the statistical performance of the predicted site-specific EONR using the validation data set. Detailed descriptions of these steps are provided below. The *R* and *python* codes that implement the simulation analyses are available at https://github.com/tmieno2/ML_VRA.git.

@@

### Underlying Yield Response Function

We modeled the yield response function as having the following Mitscherlich-Baule functional form, which is flexible in its shape, allows for the existence of a yield plateau, and has been shown to be a plausible specification for N fertilization models for corn [@frank1990; @llewelyn1997; @paris1992].

\begin{equation}
f(N, \mathbf{c}) = ymax(\mathbf{c})(1-exp(\alpha(\mathbf{c}) + \beta(\mathbf{c}) N)) + \varepsilon. (\#eq:eqn5)
\end{equation}

The first term on the right-hand side of Eq. \@ref(eq:eqn5) is the deterministic yield component. The variable $N$ represents the nitrogen fertilizer application rate (kg/ha) which the trial designs varied by plot. Also in Eq. \@ref(eq:eqn3), $\alpha$, $\beta$, and $ymax$ are functions of a vector of field characteristics variables $\mathbf{c}=(c_1,...,c_K)$, which vary spatially on the field. For example, $\alpha$ and $ymax$, which affect yield independently from the applied N fertilizer rate, might be interpreted as depending on the ingredient soil N content supplied by residual N, soil organic matter, etc. $\beta$ can be interpreted as reflecting how efficiently the crop uses N to create grain mass, and might depend on soil properties that affect the availability of applied N fertilizer to the crop, say due to N immobilization, leaching, denitrification, and volatilization [@johnson2005nitrogen; @alva2006nitrogen]. The error term of the yield $\varepsilon$ represents the composite of all the unobserved factors that affect crop yield beyond $\alpha$, $\beta$, $ymax$ and $N$.

A fictional "researcher" is modeled as not knowing the function form of $f$, as having limited knowledge about which variables make up $\mathbf{c}$, or even the number of elements of $\mathbf{c}$, and as having limited ability to measure the values that many of the variables that constitute $\mathbf{c}$ take on across the field.

@@

Let $i \in \{1,2,...,1536\}$ index a simulated field's subplots and $j \in \{1,2,...,36\}$ index cells within subplots. Let $N^{i}$ be the N fertilization rate on plot $i$, and let $\mathbf{c}^{i,j}$ be the value of $\mathbf{c}$ on cell $(i,j)$, Eq. \@ref(eq:eqn6) shows the yield level on cell $j$ in subplot $i$ is affected by the N rate applied on subplot $i$:

\begin{equation}
f(N^{i}, \mathbf{c}^{i,j}) = ymax(\mathbf{c}^{i,j})(1-exp(\alpha(\mathbf{c}^{i,j}) + \beta(\mathbf{c}^{i,j}) N^{i})) + \varepsilon^{i,j}. (\#eq:eqn6)
\end{equation}

To simplify notation, denote  $\alpha(\mathbf{c}^{i,j})$ as $\alpha^{i,j}$, $\beta(\mathbf{c}^{i,j})$ as $\beta^{i,j}$, and $ymax(\mathbf{c}^{i,j})$ as $ymax^{i,j}$. Then, a cell-specific yield function can be written as,

\begin{equation}
f^{i,j}(N) = ymax^{i,j}(1-exp(\alpha^{i,j} + \beta^{i,j} N^{i})) + \varepsilon^{i,j}. (\#eq:eqn7)
\end{equation}

Parameters $\alpha^{i,j}$, $\beta^{i,j}$, $\beta^{i,j}$, and $\varepsilon^{i,j}$ were modeled as spatially autocorrelated using unconditional Gaussian geostatistical simulation based on the spherical variogram model. The `gstat` package (version 2.0.6 [@gstat1; @gstat2]) in R was used for this purpose. Table \@ref(tab:variogram-parameters) shows the main variogram parameter values used in generating $\alpha^{i,j}$, $\beta^{i,j}$, $ymax^{i,j}$, and $\varepsilon^{i,j}$, which were chosen to generate yields consistent with those observed in experiments conducted by the DIFM project. (Codes used to generate parameter values are available at https://github.com/tmieno2/ML_VRA.git.) The range and nugget of all parameters were assumed to be $400$ m and 0. The means of $ymax$, $\alpha$, $\beta$, and $\varepsilon$ were assumed to be $12000$ kg/ha, $-0.5$, $0$ and $0$, and their sills were assumed to be $2.0 \times 10^{6}$ kg/ha, $0.02$, $1.0$ and $0.015$. Figure \@ref(fig:fig-parameters) shows an example of the generated parameters.

```{r fig-parameters, fig.width=6, fig.height=5, fig.cap = "An Illustrative Example of Spatial Distributions of Field Characteristics", cache = TRUE}
# field_alpha + field_beta + field_ymax
grid.arrange(field_ymax, field_alpha, field_beta, field_m_error, ncol = 2, nrow = 2)
```

Each cell $(i,j)$'s "true" optimal N application rate, denoted $N_{opt}^{i,j}$, was calculated by solving a profit maximization problem:

\begin{equation}
N_{opt}^{i,j} = \operatorname*{argmax}_{N}(P_{C} f^{i,j}(N) - P_N N), (\#eq:eqn8)
\end{equation}

where, $p =\$0.138/kg$ and $w =\$1.323/kg$ were the assumed prices of the corn and N fertilizer. Figure \@ref(fig:fig-optN) maps an example of the identified "true" cell-specific optimal N rates.

```{r fig-optN, fig.width=3, fig.height=4, fig.cap = "An illustrative example of a spatial distribution of true optimal N rate"}
vis_optN_cell
```

### Trial Design

In each simulation round, five field-trial N rates were determined based on the range of the round's cell-specific EONRs. 
Let $Q_q$ ($q \in [0,1]$) denote the $100 \times q$ percent quantile of the round's cell-specific EONRs. Then, the five trial design rates were $N_1 = Q_{0.05} - 20$, $N_5 = Q_{0.95} + 20$, $N_2 = N_1 + 0.25(N_5 - N_1)$, $N_3 = N_1 + 0.50(N_5 - N_1)$, $N_4 = N_1 + 0.75(N_5 - N_1)$. 
In each round, a Latin square design was followed to assign each plot one of the five N rates so established. Figure \@ref(fig:fig-nExp) shows an example of a trial design. The N rate was orthogonal to the error term in every simulation's example of trial design. This is important because if they were correlated, bias in the estimation of the impact of N on yield would be introduced into all of the ML methods. Since the treatment effect at each leaf is obtained simply by taking the difference of the average value of the dependent variable between the control and treated group within the leaf, CF requires the exogeneity of the treatment conditional on the independent variables to estimate the treatment effect consistently. Note that the violation of this condition would lead to biased estimation using the CF and the other ML methods.

```{r fig-nExp, fig.width=3, fig.height=4, fig.cap = "An illustrative example of experimental design of N"}
field_Ndesign
```

### Yield Data Generation and Aggregation

A cell-level data set for the whole field, $\{(y^{i,j}, N_{opt}^{i,j}, N^{i,j}, \mathbf{c}^{i,j}): i \in \{1,2,...,1536\}, j \in \{1,2,...,36\}\}$ was generated. For a generic cell $(i,j)$, yield $y^{i,j}$ was generated through Eq. \@ref(eq:eqn5), applying parameter values $\mathbf{c^{i,j}}=(\alpha^{i,j}, \beta^{i,j}, ymax^{i,j})$, the disturbance term level $\varepsilon^{i,j}$, and the cell's assigned experimental N rate, $N^{i,j}$. Figure \@ref(fig:field-yield-cell) presents the yield map from one of the simulations.

```{r field-yield-cell, fig.width=3, fig.height=4, fig.cap = "An illustrative example of a yield distribution map at the cell resolution"}
vis_yield_cell
```

Data were aggregated up to the subplot level by taking the mean of the cell-level values inside each subplot, which were the unit of observation. The subplot data were therefore

\begin{equation}
\{(y^{i}, N_{opt}^{i}, N^{i}, \mathbf{c}^{i}) = \frac{1}{36}\sum_{j=1}^{36} (y^{i,j}, N_{opt}^{i,j}, N^{i,j}, \mathbf{c}^{i,j}): i \in \{1,2,...,1536\}\}. (\#eq:eqn9)
\end{equation}

Figure \@ref(fig:field-yield-subplot) shows an example of observed yield data. Note that data generation at the cell level was conducted to reflect the real-world spatial heterogeneity in soil/field characteristics within the subplot level. But data aggregation to the subplot level was conducted because the large sizes of harvesters make it impossible to monitor yields accurately at smaller scales.

```{r field-yield-subplot, fig.width=3, fig.height=4, fig.cap = "An illustrative example of a yield distribution map at the subplot resolution"}
vis_yield_subplot
```

@@

To evaluate the performances of RF, BRF, CNN and CF, we examined four modeling scenarios, each defined by a dataset assumed to be available to the researcher. The first modeling scenario, denoted "aby" assumed that the researcher has a dataset of the actual values $\alpha^{i,j}$, $\beta{i,j}$, $ymax^{i,j}$ for every cell $(i,j)$ and use $\alpha$, $\beta$ and $ymax$ as covariates in the ML models. This scenario represents an ideal situation for the researcher, in which all the relevant variables are observed, but the true functional form of heterogeneous yield response functions are not known. More realistic scenarios can be compared with this ideal scenario.

In subsequent modeling scenarios, we made the simulated experimental data more realistic by varying the degree to which the researcher understood what the needed covariates' values were those in the data set. In the second scenario, denoted "abytt," we introduced additional variables $\theta_1$ and $\theta_2$, and the researcher's data set contained values of $\alpha^{i}$, $\beta^{i}$, $ymax^{i}$, $\theta_1^{i}$ and $\theta_2^{i}$ for all subplots $i=1,...,1536$. $\theta_1$ and $\theta_2$ had no effect on yield, but their cell-level values were spatially correlated with $\beta^{i,j}$, and so their subplot-level values were correlated with $\beta^{i}$. This scenario was meant to reflect the lack of scientific consensus about which variables explain heterogeneous yield response to N, making it likely that the researcher would include irrelevant variables like $\theta_1$ and $\theta_2$ in real-world data analysis.

In the third scenario, denoted "aabbyy," the researcher's data set did not contain values of $\alpha^{i}$, $\beta^{i}$, or $ymax^{i}$. Rather, it contained subplot-level values of permutations of $\alpha$, $\beta$, and $ymax$. Specifically, for each cell $(i,j)$, we assigned one of a set of spatially autocorrelated weights $r_{\alpha}^{i,j} \in (0,1)$ to the $\alpha$ variable, another set of spatially autocorrelated weights $r_{\beta}^{i,j} \in (0,1)$ to the $\beta$ variable, and another set of spatially autocorrelated weights $r_{ymax}^{i,j} \in (0,1)$ to the ${ymax}^{i,j}$ variable. (There was no spatial correlation between the $r_{\alpha}^{i,j}$ and $r_{\beta}^{i,j}$ data, between $r_{\alpha}^{i,j}$ and $r_{ymax}^{i,j}$ data, or between $r_{\beta}^{i,j}$ and $r_{ymax}^{i,j}$.) Six new covariates were created with the weights and the original covariates $\alpha^{i,j}$, $\beta^{i,j}$ and $ymax^{i,j}$. 
These were 
$\alpha_{1}^{i,j} = r_{\alpha}^{i,j} \alpha^{i,j}$, $\alpha_{2}^{i,j} = (1- r_{\alpha}^{i,j}) \alpha^{i,j}$,
$\beta_{1}^{i,j} = r_{\beta}^{i,j} \beta^{i,j}$, $\beta_{2}^{i,j} = (1- r_{\beta}^{i,j}) \beta^{i,j}$,
$ymax_{1}^{i,j} = r_{ymax}^{i,j} ymax^{i,j}$, and $ymax_{2}^{i,j} = (1- r_{ymax}^{i,j}) ymax^{i,j}$.
For each $\alpha$, $\beta$ and $ymax$, these procedures created two new covariates that were highly spatially correlated with the original covariate, which was meant to reflect that many of the soil/field characteristics used to analyze OFPE data are spatially correlated. The values of $\alpha^{i,j}$, $\beta^{i,j}$ and $ymax^{i,j}$ did not indicate specific soil or field characteristics, but were functions of such. In reality, scientists cannot directly observe variables that can accurately predict yield plateau level, soil N content and N uptake efficiency. Rather, they use multiple observed soil/field characteristics to explain such phenological phenomena. For example, instead of using $ymax^{i,j}$ directly, some topographical land features and soil properties such as elevation and soil sand content could be used as yield-limiting factors [@jiang2004effect; @kravchenko2000correlation]. In reality, scientists may include more than three variables as covariates. This scenario should reduce the accuracy of EONR modeling for all the models compared to the ideal case.

@@

The RF, BRF, and CNN methods all followed the same conceptual steps to estimate site-specific EONR: 1) estimate the yield production functions and 2) calculate site-specific EONRs based on that estimated function. For the given modeling scenario, RF and BRF predict yield by using all the available covariates as explanatory variables in the estimation process. The `grf` package (version 1.2.0 [@tibshirani2018package]) in R was used for RF and BRF modeling. RF, BRF, and CF from the `grf` package have several tunable hyperparameters (e.g., the minimum node size in each tree, the number of covariates used for node splitting, and parameters involving honest tree-building process), and the optimal values for these parameters were selected by cross-validation. Individual trees in a forest were built with randomly subsampled observations, leading to a different prediction in every forest even when the same data set was used. To reduce the variance of predictions for prediction accuracy, we set the number of trees (*num.trees*) at a high value, $4000$.

@@

Unlike RF, BRF, and CNN, CF estimates the impact of a binary treatment. In our context, CF estimates changes in yields (i.e., $\hat{\tau}_{N^{con} \rightarrow N^{tre}}(\mathbf{\Omega^{i}})$) caused by changes in N application rates from one experimental rate (control N rate, i.e., $N^{con}$) to another experimental rate (treatment N rate, i.e., $N^{tre}$). Since we have five N application rates, four "experiments" (treatments) are identified based on four ($N^{con}$, $N^{tre}$) combinations. A possible grouping to make such pairwise N application rate combinations, which we call the "CF-base," used $(N_1, N_2),(N_1, N_3),(N_1, N_4)$, and $(N_1, N_5)$. In this approach, the treatment effect is always estimated against the lowest N rate ($N_1$) as $N^{con}$.

@@

### Model Evaluation

The accuracy of EONR estimation was judged based RMSE ($kg/ha$) of the EONR estimation against the true EONR at the subplot level. Using the EONR estimates, we also calculated profit-deficits ($\$/ha$) (i.e., $\hat{\pi}_{def}$) we calculated relative to the true maximum profit at the subplot level, defined as profit under the true yield response functions evaluated at $N_{opt}^{i}$. For RF, BRF, and CNN, the accuracy of yield prediction was judged based on the RMSE ($kg/ha$) of the yield prediction against the true yields at the subplot level.

@@

## EONR Estimation

@@

## The relationship between yield prediction performance and EONR prediction performance
For RF, BRF and CNN, Figure \@ref(fig:plot-y-optN) plots the RMSE values of EONR estimation against the RMSE values of yield prediction, and the red lines show the best-fit line describing the relationship between them. For all the models the regression line was almost flat, showing little if any correlation between the two measures. In other words, the quality of a model's yield predictions provided little if any information about the quality of its EONR predictions.

```{r plot-y-optN, fig.width=7, fig.height=5, fig.cap = "The relationship in the performance between EONR estimation and yield prediction by ML methods and modeling scenarios"}
fig_y_optN
```

@@

@@

# (APPENDIX) Appendix {-}

# Appendix A: Causal Tree Example {-}

Figure \@ref(fig:ct-ex) shows an example of CT where the treatment is an increase in seed rate from $27,000$ seeds/ha (control) to $31,000$ seeds/ha (treated) with apparent electrical conductivity ($\textit{ecs}$)($dS/m$) and topographical $\textit{slope}$ ($degree$) as variables to explain the heterogeneity of the impact of the treatment impact. In the first node, all the samples ($1024$ observations) belong to the node, and the causal effect of the treatment was calculated as the difference in the mean yield of the treated and control groups, which turned out to be $593$ kg/ha in this example.

```{r ct-ex, fig.width=4, fig.height=4, fig.align="center", fig.cap = "An illustrative example of causal tree"}
fancyRpartPlot(opfit, sub = "")
```

`r ftext('NOTE: The first number in each node indicates the predicted treatment effects (in kg/ha). The number and the percentage of observations belonging to the node are shown below that.', fp_text(font.size = 9))`

The first data split criterion is whether the $\textit{ecs}$ value is greater or less than $25$. The observations with $\textit{ecs}$ greater than or equal to $25$ move to the left node, while those with $\textit{ecs}$ less than $25$ move to the right. For each of the subsamples, the treatment effect is identified as the difference in the mean yields of the treated and control groups within the subsample. The treatment effect for the low-$\textit{ecs}$ observations was $619$ kg/ha, and for the high-$\textit{ecs}$ observations was $982$ kg/ha. This means that for the impact of the increase in seed rate the treatment effect is heterogeneous over space, having a larger positive impact on yield in high-$\textit{ecs}$ areas. Trees are further developed to have more groups of subsamples to allow for more flexible representations of the heterogeneous treatment effects. The estimated heterogeneous treatments effects can then be used to identify site-specific EOIRs. It is noteworth that at no point CF estimates yield "level".

>>>>

```{r echo = F, include = FALSE, cache = F}
library(knitr)
library(here)
here::i_am("GitControlled/Writing/manuscript_cea.rmd")
# opts_knit$set(root.dir = here())

@@

opts_chunk$set(
  fig.align = "center",
  fig.retina = 5,
  warning = F,
  message = F,
  cache = TRUE,
  echo = F
)
```

```{r, cache = FALSE}
# === packages ===#
library(tidyverse)
library(data.table)
library(sf)
library(png)
library(RColorBrewer)
library(patchwork)
library(plotly)
library(magrittr)
library(ggplot2)
library(viridis)
library(tmap)
library(ggthemes)
library(mgcv)
library(ggpubr)
library(flextable)
library(officer)
library(officedown)
library(modelsummary)
library(raster)
library(bookdown)
library(png)
library(grid)
library(causalTree)
library(rpart)
library(rattle)
```

@@

CF was developed specifically for identifying heterogeneous treatment effects [@athey2016recursive; @Wager2018a]. Like RF, CF works by building an ensemble of trees, called *causal trees* (CT), by recursively partitioning covariate space into two leaves, where each leaf estimates conditional average treatment effects as the mean differences between the dependent variable (here, yield) of the treated and control observations that belong to the leaf (see Appendix \@ref(ctree-ex) for an example of a causal tree that can help further clarify how CF works).

@@

Monte Carlo simulations were conducted for nitrogen experiments for corn production to estimate site-specific EONRs. The simulations' first step generated one thousand synthetic fields that "resembled" an actual field. Field characteristics and yield were simulated assuming the Mitscherlich-Baule (MB) yield response function. We used the RF, BRF, and CNN ML methods to estimate yield response functions, and we used the CF method to estimate treatment effects. For each method, we estimated site-specific EONRs using the training data set, and checked the statistical performance of the predicted site-specific EONR using the validation data set. Detailed descriptions of these steps are provided below. The *R* and *python* codes that implement the simulation analyses are available at https://github.com/tmieno2/ML_VRA.git.

@@

@@

We modeled the yield response function as having the Mitscherlich-Baule functional form, which is flexible in its shape, allows for the existence of a yield plateau, and has been shown to be a plausible specification for N fertilization models for corn [@frank1990; @llewelyn1997; @paris1992].

Let $i \in \{1,2,\dots,1536\}$ index a simulated field's subplots and $j \in \{1,2,\dots,36\}$ index cells within subplots. Let $N_{i}$ be the N fertilization rate on subplot $i$. Eq. \@ref(eq:eqn6) shows the yield level on cell $j$ in subplot $i$:

\begin{equation}
f(N_{i}: \alpha_{i,j}, \beta_{i,j}, ymax_{i,j}) = ymax_{i,j}(1-exp(\alpha_{i,j} + \beta_{i,j} \cdot N_{i})) + \varepsilon^{i,j}. (\#eq:eqn6)
\end{equation}

The first term on the right-hand side of Eq. \@ref(eq:eqn6) is the deterministic yield component. The variable $N$ represents the nitrogen fertilizer application rate (kg/ha) which the trial designs varied by plot. Also in Eq. \@ref(eq:eqn6), $\alpha$, $\beta$, and $ymax$ are spatially varying parameters that govern how yield level is determined in conjunction with the applied N rate. They reflect local cell-level production conditions, such as soil characteristics and topography. For example, $\alpha$ and $ymax$, which affect yield independently from the applied N fertilizer rate, might be interpreted as depending on the ingredient soil N content supplied by residual N, soil organic matter, etc. $\beta$ can be interpreted as reflecting how efficiently the crop uses N to create grain mass, and might depend on soil properties that affect the availability of applied N fertilizer to the crop, say due to N immobilization, leaching, denitrification, and volatilization [@johnson2005nitrogen; @alva2006nitrogen]. The error term of the yield $\varepsilon$ represents the composite of all the unobserved factors that affect crop yield beyond $\alpha$, $\beta$, $ymax$ and $N$. Each cell $(i,j)$'s "true" optimal N application rate, denoted $N^{opt}_{i,j}$, was calculated by solving a profit maximization problem by cell assuming that the prices of corn and N fertilizer were $\$0.138/kg$ and $\$1.323/kg$, respectively. These prices are used consistently for our simualtions. See Appendix \@ref(par-gen) for further details on how the paremeters were generated and sample maps of the parameters and true economically optimal N rates.

### Trial Design and Yield Data Generation for Modeling

In each simulation round, five field-trial N rates were determined based on the range of the round's cell-specific EONRs. 
Let $Q_q$ ($q \in [0,1]$) denote the $100 \times q$ percent quantile of the round's cell-specific EONRs. Then, the five trial design rates were $N_1 = Q_{0.05} - 20$, $N_5 = Q_{0.95} + 20$, $N_2 = N_1 + 0.25(N_5 - N_1)$, $N_3 = N_1 + 0.50(N_5 - N_1)$, $N_4 = N_1 + 0.75(N_5 - N_1)$. 
In each round, a Latin square design was followed to assign each plot one of the five N rates so established. The left-most panel of Figure \@ref(fig:fig-nExp-yield) shows an example of a trial design. The N rate was orthogonal to the error term in every simulation's example of trial design. This is important because if they were correlated, bias in the estimation of the impact of N on yield would be introduced into all of the ML methods. Since the treatment effect at each leaf is obtained simply by taking the difference of the average value of the dependent variable between the control and treated group within the leaf, CF requires the exogeneity of the treatment conditional on the independent variables to estimate the treatment effect consistently. Note that the violation of this condition would lead to biased estimation using the CF and the other ML methods.

A cell-level data set for the whole field, $\{(y_{i,j}, N^{opt}_{i,j}, N_{i,j}, \mathbf{c}_{i,j}): i \in \{1,2,\dots,1536\}, j \in \{1,2,\dots,36\}\}$ was generated by plugging in the parameter values, the error term, and the assigned nitrogen rates into \@ref(eq:eqn6). The cell-level data were then aggregated up to the subplot level by taking the mean of the cell-level values inside each subplot, which were the unit of observation. The subplot data were therefore

\begin{equation}
\{(y_{i}, N^{opt}_{i}, N_{i}) = \frac{1}{36}\sum_{j=1}_{36} (y_{i,j}, N^{opt}_{i,j}, N_{i,j}): i \in \{1,2,\dots,1536\}\}. (\#eq:eqn9)
\end{equation}

The right panel of Figure \@ref(fig:fig-nExp-yield) shows an example of observed yield data. Note that data generation at the cell level was conducted to reflect the real-world spatial heterogeneity in soil/field characteristics within the subplot level. But data aggregation to the subplot level was conducted because the large sizes of harvesters make it impossible to monitor yields accurately at smaller scales.

```{r fig-nExp-yield, fig.width=7, fig.height=6, fig.cap = "An example of experimental design of N and generated yield at the subplot resolution"}

(field_Ndesign + ggtitle("(a) Trial Design")) |
  (vis_yield_subplot + ggtitle("(b) Simulated Yield Level"))
```

@@

To evaluate the performances of RF, BRF, CNN and CF, we examined four modeling scenarios, each defined by a dataset assumed to be available to the researcher. The first modeling scenario, denoted "aby" assumed that the researcher has a dataset of the actual values $\alpha_{i}$, $\beta_{i}$, $ymax_{i}$ in the yield response function for every subplot $i$ and use $\alpha$, $\beta$ and $ymax$ as covariates in the ML models. This scenario represents an ideal situation for the researcher, in which all the relevant variables are observed, but the true functional form of heterogeneous yield response functions are not known. More realistic scenarios can be compared with this ideal scenario.

In subsequent modeling scenarios, we made the simulated experimental data more realistic by varying the degree to which the researcher understood what the needed covariates' values were those in the data set. In the second scenario, denoted "abytt," we introduced additional variables $\theta^1$ and $\theta^2$, and the researcher's data set contained values of $\alpha_{i}$, $\beta_{i}$, $ymax_{i}$, $\theta^1_{i}$ and $\theta^2_{i}$ for all subplots $i=1,\dots,1536$. $\theta^1$ and $\theta^2$ had no effect on yield, but they were spatially correlated with $\beta^{i}. This scenario was meant to reflect the lack of scientific consensus about which variables explain heterogeneous yield response to N, making it likely that the researcher would include irrelevant variables like $\theta^1$ and $\theta^2$ in real-world data analysis.

In the third scenario, denoted "aabbyy," the researcher's data set did not contain values of $\alpha_{i}$, $\beta_{i}$, or $ymax_{i}$. Rather, it contained subplot-level values of permutations of $\alpha$, $\beta$, and $ymax$. Each of the three parameters were split into two parts in a spatially correlated manner, and six new covariates were created (see Appendix \@ref(split-pars) for the details about this prameter splitting process). These were $\alpha^{1}_{i}$, $\alpha^{2}_{i}$, $\beta^{1}_{i}$, $\beta^{2}_{i}$, $ymax^{1}_{i}$, and $ymax^{2}_{i}$. The two splitted variables were spatially correlated to reflect that many of the soil/field characteristics used to analyze OFPE data are spatially correlated. The values of $\alpha_{i}$, $\beta_{i}$ and $ymax_{i}$ did not indicate specific soil or field characteristics, but were functions of such. In reality, scientists cannot directly observe variables that can accurately predict yield plateau level, soil N content and N uptake efficiency. Rather, they use multiple observed soil/field characteristics to explain such phenological phenomena. For example, instead of using $ymax_{i}$ directly, some topographical land features and soil properties such as elevation and soil sand content could be used as yield-limiting factors [@jiang2004effect; @kravchenko2000correlation]. In reality, scientists may include more than three variables as covariates. This scenario should reduce the accuracy of EONR modeling for all the models compared to the ideal case.

@@

The RF, BRF, and CNN methods all followed the same conceptual steps to estimate site-specific EONR: 1) estimate the yield production functions and 2) calculate site-specific EONRs based on that estimated function. For the given modeling scenario, RF and BRF predict yield by using all the available covariates as explanatory variables in the estimation process. The `grf` package (version 1.2.0 [@tibshirani2018package]) in R was used for RF and BRF modeling. RF ann BRF from the `grf` package have several tunable hyperparameters (e.g., the minimum node size in each tree, the number of covariates used for node splitting, and parameters involving honest tree-building process), and the optimal values for these parameters were selected by cross-validation. Individual trees in a forest were built with randomly subsampled observations, leading to a different prediction in every forest even when the same data set was used. To reduce the variance of predictions for prediction accuracy, we set the number of trees (*num.trees*) at a high value, $4000$.

@@

Unlike RF, BRF, and CNN, CF estimates the impact of a binary treatment. In our context, CF estimates changes in yields (i.e., $\hat{\tau}_{N^{con} \rightarrow N^{tre}}(\mathbf{\Omega^{i}})$) caused by changes in N application rates from one experimental rate (control N rate, i.e., $N^{con}$) to another experimental rate (treatment N rate, i.e., $N^{tre}$). Since we have five N application rates, four "experiments" (treatments) are identified based on four ($N^{con}$, $N^{tre}$) combinations. A possible grouping to make such pairwise N application rate combinations, which we call the "CF-base," used $(N_1, N_2),(N_1, N_3),(N_1, N_4)$, and $(N_1, N_5)$. In this approach, the treatment effect is always estimated against the lowest N rate ($N_1$) as $N^{con}$. The `grf` package (version 1.2.0 [@tibshirani2018package]) in R was used for CF estimation. Just like RF and BRF, hyperparameters were tuned using cross-validatiaon.

@@

### Evaluation of yield and EONR predictions

For evaluation of yield and EONR predictions, a dataset from a different simulation round was used as the test dataset. For example, for simulation number 1, train dataset from simulation number 2 was used as the test dataset. This process works because the train datasets were generated by the same data generating process, yet have different underlying field parameters ($\alpha$, $\beta$, $ymax$). Furtehr, each of the "train" datasets are generated independet of one another. Thefore, for a particular simulation round, the dataset in any of the other simualtions can serve as a valid test dataset. This practice avoid creating another set of 1000 simulations fields unnecessarily.

The accuracy of EONR estimation was judged based on RMSE ($kg/ha$) of the EONR estimation against the true EONR at the subplot level. Using the EONR estimates, we also calculated profit-deficits ($\$/ha$) (i.e., $\hat{\pi}_{def}$) we calculated relative to the true maximum profit at the subplot level, defined as profit under the true yield response functions evaluated at $N^{opt}^{i}$. For RF, BRF, and CNN, the accuracy of yield prediction was judged based on the RMSE ($kg/ha$) of the yield prediction against the true yields at the subplot level.

@@

## Comparison of EONR estimation accuracy

@@

## The relationship between yield and EONR prediction performances.

For RF, BRF and CNN, Figure \@ref(fig:plot-y-optN) plots the RMSE values of EONR estimation against the RMSE values of yield prediction, and the red lines show the best-fit line describing the relationship between them. For all the models the regression line was almost flat, showing little if any correlation between the two measures. In other words, the quality of a model's yield predictions provided little if any information about the quality of its EONR predictions.

```{r plot-y-optN, fig.width=7, fig.height=7, fig.cap = "The relationship in the performance between EONR estimation and yield prediction by ML methods and modeling scenarios"}
fig_y_optN
```

@@

# Limitations and Future Studies

@@

# Supplementary Appendix {-}

```{r, child = "appendix.Rmd"}
```
